---
title: "Take-home Exercise 2: Application of Spatial and Spatio-temporal Analysis Methods to Discover the Distribution of Dengue Fever in Tainan City, Taiwan - Part 2"
format: 
  html: 
    code-fold: true
    code-summary: "Show the code"
    code-overflow: wrap
    code-line-numbers: true
    code-copy: hover
execute: 
  eval: true
  echo: true
  freeze: auto
---

```{r}
pacman::p_load(sf, tidyverse, tmap, sfdep)
```

Load data in RDS format

```{r}
tainan_dengue <- read_rds("../../data/rds/tainan_dengue.rds")
tainan_dengue_s3 <- complete_spacetime_cube(tainan_dengue)
```

# Global Spatial Autocorrelation Analysis

This exercise expected the `sfdep` methods taught in [In-class Exercise 5]("In-class_Ex/In-class_Ex05/In-class_Ex05.html") to derive the various statistic values and explain them.

## Derive contiguity weights with Queen's method

Get neighbour list

-   `activate` will retrieve geometry column back (as spacetime object doesn't show geomotry)

```{r}

```

# Local Spatial Autocorrelation Analysis


# Emerging Hotspot Analysis