---
title: "Take-home Exercise 2: Application of Spatial and Spatio-temporal Analysis Methods to Discover the Distribution of Dengue Fever in Tainan City, Taiwan - Geospatial Data Wrangling"
format: 
  html: 
    code-fold: true
    code-summary: "Show the code"
    code-overflow: wrap
    code-line-numbers: true
    code-copy: hover
    self-contained: true
execute: 
  eval: true
  echo: true
  freeze: auto
---

```{r}
pacman::p_load(sf, tidyverse, tmap, sfdep, lubridate)
```

# Import Data

The following datasets are used in this exercise.

-   `TAIWAN_VILLAGE_2020`, a geospatial data of village boundary of Taiwan. It is in ESRI shapefile format. The data is in Taiwan Geographic Coordinate System TWD97.

-   `Dengue_Daily.csv`, an aspatial data of reported dengue cases in Taiwan since 1998.

```{r}
tainan <- st_read(dsn = "../../data/geospatial",
                  layer = "TAINAN_VILLAGE")

dengue <- read_csv("../../data/aspatial/Dengue_Daily.csv")
```

# Preparing data for analysis

## Study area preparation

According to the task objective given, this exercise requested for towns labelled D01, D02, D04, D06, D07, D08, D32 and D39

```{r}
tainan <- tainan[tainan$TOWNID %in% c("D01", "D02","D04","D06","D07","D08","D32","D39"),]
plot(tainan)
```
After checking `VILLNAME` available in `tainan`, there are repeated village names that are under different towns. Assign new names to each village by combining existing town name with current village name

```{r}
#| eval: false
#| echo: false
table(tainan$VILLNAME)
```

```{r}
tainan$VILLNAME <- paste(tainan$TOWNNAME, tainan$VILLNAME,
                         sep = "_")
```


```{r}
#| eval: false
#| echo: false
write_rds(tainan, "../../data/rds/tainan.rds")
```

## Dengue fever layer preparation

This exercise requires dengue cases in Tainan, Taiwan. The original tibble is filtered by the condition where county name is Tainan and the desired time period of week 31 to 50 of 2023, and saved back to the same variable name.

```{r}
dengue <- dengue %>% 
  mutate(年 = factor(year(發病日)),
         week = epiweek(發病日),
         VILLNAME = paste(居住鄉鎮, 居住村里, sep = "_")) %>% 
  filter(年 == 2023,
         居住縣市 == "台南市",
         week >= 31,
         week <= 50) 
```

Convert `dengue` to `sf` object such that the object would be able to be projected to `TWD97`. Before conversion, check that there is no missing value in coordinates

```{r}
sum(dengue$最小統計區中心點X == "None")
sum(dengue$最小統計區中心點Y == "None")
```

Drop those with None values: Currently could not find a effective way to fill in, so to not obstruct the conversion into a `sf` object, I decided to drop the rows with "None" values for the coordinates columns. Also, to standardise the village name for `dengue` dataset.

```{r}
dengue <- dengue %>% 
  filter(最小統計區中心點X != "None",
         最小統計區中心點Y != "None",
         VILLNAME  %in% tainan$VILLNAME)
sum(dengue$最小統計區中心點X == "None")
sum(dengue$最小統計區中心點Y == "None")
```

Convert `dengue` to `sf` object and assign projected coordinate system

```{r}
dengue <- st_as_sf(dengue,
                   coords = c("最小統計區中心點X","最小統計區中心點Y"),
                   crs = 3824)
```


```{r}
#| echo: false
#| eval: false
write_rds(dengue, "../../data/rds/dengue.rds")
```

# Get derived dengue layer

Using `spacetime()` function to create `spacetime` object

```{r}
colnames(dengue)
```

```{r}
dengue_grp <- dengue %>%
  group_by(VILLNAME, week) %>%
  summarise(num_dengue_cases = n()) %>% 
  filter(VILLNAME != "None") %>% 
  st_drop_geometry()

tainan <- tainan %>% 
  filter(VILLNAME %in% unique(dengue_grp$VILLNAME))
```

As the time-series is incomplete, stating only weeks that have recorded instances of dengue, weeks with `num_dengue_cases = 0` needs to be included as well.

```{r}

for (vill in dengue_grp$VILLNAME) {
  vill_record <- dengue_grp %>%
    filter(VILLNAME == vill)
  for (wk in 31:50) {
    if (!(wk %in% vill_record$week)) {
      dengue_grp[nrow(dengue_grp)+1,] = list(vill, wk, 0)
    }
  }
}
```


```{r}
tainan_dengue_st <- spacetime(.data = dengue_grp,
                              .geometry = tainan,
                              .loc_col = "VILLNAME",
                              .time_col = "week")
is_spacetime_cube(tainan_dengue_st)
```

The derived `tainan_dengue` layer is then saved as a RDS file.
```{r}
#| eval: false
#| echo: false
write_rds(tainan_dengue_st,"../../data/rds/tainan_dengue.rds")
```

Please go to [Report](Take-home_Ex/Take-home_Ex02/Take-home_Ex02_Part2.html) for further details.
