---
title: "Take-home Exercise 2: Application of Spatial and Spatio-temporal Analysis Methods to Discover the Distribution of Dengue Fever in Tainan City, Taiwan - Part 2"
format: 
  html: 
    code-fold: true
    code-summary: "Show the code"
    code-overflow: wrap
    code-line-numbers: true
    code-copy: hover
execute: 
  eval: true
  echo: true
  freeze: auto
---

```{r}
pacman::p_load(sf, tidyverse, tmap, sfdep)
```

Load data in RDS format

```{r}
dengue <- readRDS("../../data/rds/dengue.rds")
tainan <- readRDS("../../data/rds/tainan.rds")
```


## Get the derived dengue layer

```{r}
tmap_mode("view")
tm_shape(dengue)+
  tm_dots()
```

Derive contiguity weight with Queen's method

```{r}
colnames(dengue)
```

Find the counts of dengue cases in each town within the county

-   The results showed that there are 344 villages with recorded cases in `dengue_grp` dataset, with a huge 17,945 cases without village name.

I then matched the `tainan` dataset with the `dengue_grp` dataset, left joined using the village names that is presented in both datasets.

```{r}
dengue_grp <- dengue %>%
  group_by(感染村里) %>%
  summarise(num_dengue_cases=n()) %>%
  st_drop_geometry()

dengue_grp

tainan_dengue <- left_join(tainan, dengue_grp,
                           by = join_by(VILLNAME == 感染村里))

tainan_dengue$num_dengue_cases <- tainan_dengue$num_dengue_cases %>%
  replace_na(0)
tainan_dengue
```


