---
title: "Hands-on Exercise 01: Geospatial Data Wrangling with R"
id: "remove 1"
---

*Do note to download all files from the given data sources stated.*

With reference to this [hands-on exercise](https://r4gdsa.netlify.app/chap01.html)

# Getting Started

### Installing and Loading packages and dataset

In this section, I will be installing and load sf and tidyverse R packages using pacman

```{r}
# install.packages("pacman")
pacman::p_load(sf,tidyverse)
```

Importing the geospatial data files using st_read() *Full file path works*

```{r}
mpsz <- st_read("/Users/tangtang/Desktop/IS415 Geospatial Analytics and Applications/practice/IS415-GAA/data/geospatial/MP14_SUBZONE_WEB_PL.shp")

cycling_path <- st_read("/Users/tangtang/Desktop/IS415 Geospatial Analytics and Applications/practice/IS415-GAA/data/geospatial/CyclingPathGazette.shp")

preschool <- st_read("/Users/tangtang/Desktop/IS415 Geospatial Analytics and Applications/practice/IS415-GAA/data/geospatial/PreSchoolsLocation.kml")
```

# Checking Content of our dataset

There are many ways that we can use to retrieve summarised information on the dataset.

1.  st_geometry(dataset_name)

```{r}
st_geometry(mpsz)
```

2.  glimpse(dataset_name)

```{r}
glimpse(mpsz)
```

3.  head(dataset_name,n=5)

```{r}
head(mpsz,n=5)
```

# Plotting Geospatial data

```{r}
plot(mpsz, max.plot = 15)
plot(st_geometry(mpsz))
plot(mpsz["PLN_AREA_N"])
```

# Manipulating data

### Assigning EPSG code

This is done when coordinate system of the source data was either missing (such as due to missing .proj for ESRI shapefile) or wrongly assigned during the importing process

```{r}
st_crs(mpsz)
mpsz3414 <- st_set_crs(mpsz, 3414)
st_crs(mpsz3414)
```

### Changing coordinate system for dataset to match

This is how the preschool dataset looks like with its original coordinate system, wgs84.

```{r}
st_geometry(preschool)
```

Project transformation is done using

```{r}
preschool3414 <- st_transform(preschool,
                              crs=3414)
```

The coordinate system of preschool dataset has changed.

```{r}
st_geometry(preschool3414)
```

# Loading and Manipulating Aspatial data

```{r}
listings <- read_csv("../../data/aspatial/listings.csv")
list(listings)
```

### Coverting aspatial dataset into data frame

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
glimpse(listings_sf)
```

# Geoprocessing

```{r}
buffer_cycling <- st_buffer(cycling_path, 
                            dist=5, 
                            nQuadSegs = 30)
buffer_cycling$AREA <- st_area(buffer_cycling)
sum(buffer_cycling$AREA)
```

Question: A pre-school service group want to find out the numbers of pre-schools in each Planning Subzone.

```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
summary(mpsz3414$`PreSch Count`)
top_n(mpsz3414, 1, `PreSch Count`)
```

Question: To calculate the density of pre-school by planning subzone

```{r}
mpsz3414$Area <- mpsz3414 %>%
  st_area()
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

# EDA with R

```{r}
hist(mpsz3414$`PreSch Density`)
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
ggplot(data=mpsz3414, 
       aes(y = `PreSch Count`, 
           x= as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "Pre-school density (per km sq)",
      y = "Pre-school count")
```
