---
title: "Take-home Exercise 1: Application of Spatial Point Pattern Analysis on Grab Hailing Service in Singapore"
format: 
  html: 
    code-fold: true
    code-summary: "Show the code"
    code-overflow: wrap
    code-line-numbers: true
    code-copy: hover
    self-contained: true
execute: 
  eval: true
  echo: true
  freeze: auto
author: "Tang Xin Yi"
date: "17 Jan 2024"
---

# 1 Problem context

This exercise is looking into discovering insights relevant to the geographical and spatial distribution of Grab hailing service locations in Singapore.

In this exercise, the following packages are used.

```{r}
pacman::p_load(sf, tidyverse, tmap, spatstat, raster, maptools, spNetwork, classInt, viridis, arrow, lubridate)
```

# 2 Data

## 2.1 Background of the datasets used

The following datasets are used in this analysis exercise.

-   `GrabPosisi`: Aspatial data

-   `RoadMap`: from OpenStreetMap

-   `Master Plan 2019 Subzone Boundary (No Sea)` from Data.gov.sg

## 2.2 Import data

```{r}

grab <- read_parquet("../../data/aspatial/GrabPosisi/part-00000-8bbff892-97d2-4011-9961-703e38972569.c000.snappy.parquet")

roadmap <- st_read(dsn = "../../data/geospatial/malaysia-singapore-brunei-latest-free.shp",
                   layer = "gis_osm_roads_free_1")

mp19 <- st_read("../../data/geospatial/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")
```

From below, `grab$pingtimestamp` is not in the correct data type (it is in int class)

```{r}
head(grab)
grab$pingtimestamp = as_datetime(grab$pingtimestamp)
```

Now the `pingtimestamp` is in the correct data type `POSIXct` which represents datetime in tibble

```{r}
head(grab)
```

There are Z dimensions in `mp19` multipolygon Z, drop Z dimensions to keep the multipolygons

```{r}
mp19 <- st_zm(mp19$geometry, drop = TRUE, what = "ZM")
```

The tibble data.frame is using coordinate system `WGS 84`, but we want to use the data in `SVY21`

```{r}
st_crs(mp19)
```

```{r}
mp19 <- st_transform(mp19, crs = 3414)
```

```{r}
head(roadmap)
st_crs(roadmap)
```

`roadmap` has a similar problem as `mp19`, the coordinate system needs to be changed to `SVY 21`

```{r}
roadmap <- st_transform(roadmap, crs = 3414)
head(roadmap)
```

## 2.3 Prepare required Geographical data layers

This exercise needs the following geographical data layers

1.  Grab taxi location points either by origins or destinations

2.  Road layer within Singapore excluding outer islands

3.  Singapore boundary layer excluding outer islands

### 2.3.1 Get Grab taxi origin and destination layer

Get the origin tibble data.frame

```{r}
trip_origin <- grab %>% 
  # group by trip id
  group_by(trj_id) %>% 
  # sort: ascending by default
  arrange(pingtimestamp, .by_group = FALSE) %>% 
  # get 1st row
  filter(row_number()==1) %>% 
  mutate(weekday = wday(pingtimestamp,
                        label = TRUE,
                        abbr = TRUE),
         start_hr = factor(hour(pingtimestamp)),
         day = factor(mday(pingtimestamp)))
```

Get the destination tibble data.frame

```{r}
trip_dest <- grab %>% 
  group_by(trj_id) %>% 
  arrange(desc(pingtimestamp), .by_group = FALSE) %>% 
  filter(row_number()==1) %>% 
  mutate(weekday = wday(pingtimestamp,
                        label = TRUE,
                        abbr = TRUE),
         end_hr = factor(hour(pingtimestamp)),
         day = factor(mday(pingtimestamp)))
```

The extracted data layers are written into `rds` files for easier reference later.

```{r}
#| eval: false
#| echo: false
write_rds(trip_origin, "../../data/rds/trip_origin.rds")
write_rds(trip_dest, "../../data/rds/trip_dest.rds")
```

### 2.3.3 Get Singapore boundary layer

This part is using `mp19` and `roadmap`

Do a quick plot of the union to get the boundary layer

```{r}
plot(st_union(mp19))
```

Save the boundary layer into RDS file

```{r}
#| echo: false
#| eval: false
write_rds(st_union(mp19), "../../data/rds/sg_bount.rds")
```

### 2.3.2 Get Road layer within Singapore

Overlap the 2 dataframe to get only roads for Singapore

```{r}
road_sg <- st_intersection(roadmap, mp19)
```

Zoom in to see the roads via ID

```{r}
#| eval: false
tmap_mode("view")
tm_shape(road_sg)+
  tm_lines()
tmap_mode("plot")
```

Extract out the road layer

```{r}
#| echo: false
#| eval: false
write_rds(road_sg, "../../data/rds/road_sg.rds")
```

# 3 Kernel Density Estimation

# 4 Network-constrained Kernel Density Estimation
